---
title: "Interactive Data Visualization with Plotly"
subtitle: "3D Mathematical Surface Exploration"
author: "Rahul Vijayraghavan"
date: "January 27, 2026"
output:
  ioslides_presentation:
    widescreen: true
    smaller: false
    transition: faster
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(plotly)
```

## Introduction

**Why Plotly?**

- Interactive visualizations for the web
- Supports 2D, 3D, and statistical plots
- Seamless R integration
- No JavaScript knowledge required

**Today's Demo:**

A 3D wave interference pattern with full interactivity

## Interactive 3D Surface Plot

```{r plotly-viz, fig.width=10, fig.height=7}
# Create coordinate grid
x <- seq(-5, 5, length.out = 100)
y <- seq(-5, 5, length.out = 100)

# Generate wave interference pattern
z <- outer(x, y, function(x, y) {
  sin(sqrt(x^2 + y^2)) / sqrt(x^2 + y^2 + 1)
})

# Create interactive 3D surface
plot_ly(x = ~x, y = ~y, z = ~z) %>%
  add_surface(
    colorscale = "Viridis",
    contours = list(
      z = list(
        show = TRUE,
        usecolormap = TRUE,
        highlightcolor = "#ff0000",
        project = list(z = TRUE)
      )
    )
  ) %>%
  layout(
    title = "Wave Interference Pattern: f(x,y) = sin(r) / (r + 1)",
    scene = list(
      camera = list(
        eye = list(x = 1.87, y = 0.88, z = 0.64)
      ),
      xaxis = list(title = "X Coordinate"),
      yaxis = list(title = "Y Coordinate"),
      zaxis = list(title = "Amplitude")
    )
  )
```

## How to Interact

**Try these interactions:**

- **Rotate**: Click and drag to rotate the 3D surface
- **Zoom**: Scroll or pinch to zoom in/out
- **Pan**: Shift + drag to pan the view
- **Hover**: Hover over the surface to see coordinate values
- **Reset**: Double-click to reset the camera view

The contour lines project onto the bottom plane for better depth perception!

## Mathematical Details {.smaller}

**Function Definition:**

$$f(x, y) = \frac{\sin(\sqrt{x^2 + y^2})}{\sqrt{x^2 + y^2} + 1}$$

This creates a radial wave pattern that:

- Emanates from the origin (0, 0)
- Dampens with distance due to the denominator
- Creates concentric circular waves
- Demonstrates wave interference principles

**Technical Implementation:**

- 100 x 100 grid points (10,000 calculations)
- Viridis color scale for accessibility
- Contour projection for depth visualization

## Thank You!

**Project Information:**

- Course: Developing Data Products (Coursera)
- Date: January 27, 2026
- Created with: R Markdown + Plotly

**Resources:**

- [Plotly R Documentation](https://plotly.com/r/)
- [R Markdown Guide](https://rmarkdown.rstudio.com/)

---

Questions?
